---
title: "p8105_hw3_rh3195"
author: "Ruijie He"
date: "2023-10-14"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)
knitr::opts_chunk$set(echo = TRUE)
library(ggridges)
```

## Problem 1

```{r}
data("instacart")

instacart = 
  instacart |> 
  as_tibble()
```

> This dataset contains `r nrow(instacart)` rows, which represents the ordered items. And it contains `r nrow(instacart)` columns. Some key variables specifies the informations of each order, like `order_number`, `product_name`, `aisle`, `department`, and `order_dow`. They tells the time of the orders placed and some purchase activities for each product. In this dataset, there are `r instacart |> select(product_id) |> distinct() |> count()` products found in `r instacart |> select(user_id, order_id) |> distinct() |> count()` orders from `r instacart |> select(user_id) |> distinct() |> count()` distinct users.

**Aisle number**

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

> There are 134 aisles, and the aisle has the most items is the fresh vegetables aisle.

**Make a plot (aisle has >10000)**


```{r aisle_plot}
aisle_plot = 
  instacart %>% 
  janitor::clean_names() %>% 
  count(aisle) %>% 
  filter(n>10000) %>% 
  ggplot(aes( x= reorder(aisle, n), y = n)) +
  geom_point() +
  labs(
    title = "Number of Items",
    x = "Aisle number",
    y = "Number of item",
    caption = "Data from instacart online grocery shopping dataset 2017"
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
> The plot shows the number of items in selected aisles in an ascending order. Only aisles that contains more than 10000 items were selected.

**Three most popular items table**

```{r, message = FALSE}
popular_items =
  instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle, product_name) %>% 
  summarize(number_of_ordered_times = n()) %>% 
  arrange(aisle, -number_of_ordered_times) %>% 
  do(top_n(., 3, wt = number_of_ordered_times))

popular_items_table =
  knitr::kable(popular_items,
      col.names = c("Aisle", "Product name", "Number of ordered times"),
      title = "The 3 Most popular items in selected aisle")
```

> This table shows the three most popular item in `baking ingredients`, `dog food care`, and `packaged vegetables fruits` and lists the number of times each has been ordered.

**A 2 x 7 table**

As the `order_dow` tells the day of the week the order was placed on.

```{r}
mean_hour =
  instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) %>% 

knitr::kable(digits = 3,
             title = "Mean Hour Ordered of Selected Items")
  
mean_hour
```

> Above table shows the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. The variable `order_dow` tells the day of the week the order was placed on. The table has been `pivot_wider` to format for human readers. Pink lady apple were bought slightly earlier in the day 1, 2, 4, 6 each week than coffee ice cream. And both items were likely to be bought in the afternoon.

## Problem 2


```{r}
library(p8105.datasets)
data("brfss_smart2010")
```

**Data cleaning**

```{r}
brfss_df =
  brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health",
         response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) %>% 
  mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"), order = TRUE)) %>% 
  rename("state" = "locationabbr") %>% 
  select(topic, response, everything())

brfss_df

```


> The cleaned dataset is focusing on `topic` of "overall health" and only includes `responses` from "excellent" to "poor".

**States observed at >= 7 locations in 2002 and 2010**

```{r states 2002}
states_observed_2002 =
  brfss_df %>% 
  filter(year == 2002) %>% 
  group_by(state) %>% 
  summarize(n_observation = n_distinct(locationdesc)) %>% 
  filter(n_observation > 6) %>% 
  arrange(desc(n_observation))

states_observed_2002
```

> In 2002, the states that were observed at 7 or more locations are PA, MA, NJ, CT, NC, and FL.

```{r states 2010}
states_observed_2010 =
  brfss_df %>% 
  filter(year == 2010) %>% 
  group_by(state) %>%
  summarize(n_observation = n_distinct(locationdesc)) %>% 
  filter(n_observation > 6) %>% 
  arrange(desc(n_observation))

states_observed_2010
```

> In 2010, 14 states were observed at 7 or more locations, which are FL, NJ, TX, CA, MD, NC, NE, WA, MA, NY.


**Construct dataset limited to "excellent" response**

```{r, message = FALSE}
excellent_df =
  brfss_df %>% 
  filter(response == "Excellent") %>% 
  select(response, year, state, data_value, locationdesc) %>% 
  group_by(year, state) %>% 
  summarize(avg_value = mean(data_value))
```

*Make spaghetti plot*

```{r spaghetti_plot}
spaghetti_plot =
  excellent_df %>% 
  ggplot(aes(x = year, y = avg_value, group = state, color = state)) +
  geom_line() +
  labs(
    x = "Year",
    y = "Average value"
  ) +
  scale_color_hue ( #Describe features of colors)
    name = "States") +
  theme(legend.position = "right")

print(spaghetti_plot)
```

> The average value for states with `Excellent` response from 2002 to 2010 is between 10 and 30. This `spaghetti_plot` gives the visualization, but it is a bit crowdy as there are too many states.

**Two-panel plot show the distribution**

```{r two_panel_plot_NY}
two_panel_plot_NY =
  brfss_df %>% 
  janitor::clean_names() %>% 
  filter(year == "2010" | year == "2006", state == "NY") %>% 
  select(year, state, locationdesc, response, data_value) %>%  
  group_by(year) %>% 
  ggplot(aes(x = response, y = data_value, color = response)) +
  geom_boxplot() +
  labs(
    title = "Distrubution of data_value for responses among locations in NY on 2006 and 2010",
    x = "Response level",
    y = "Data value",
    caption = "brfss_smart2010 dataset",
  ) +
  theme(legend.position = "bottom") +
  facet_grid(. ~ year)

two_panel_plot_NY
```

> The two-panel plot visualized the distribution of `data_value` for responses from "poor" to "excellent" in 2006 and 2010 among locations in New York state. The data value for 2010 is slightly greater than in 2006.

## Problem 3

Variable `MIMS` is the MIMS values for each minute of a 24-hour day starting at midnight.

**Load and tidy two datasets**

```{r}
demo_clean =
  read.csv("./data/nhanes_covar.csv", skip = 4) %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  filter(age >= 21) %>% 
  mutate(
    sex = recode(sex, "1" = "male", "2" = "female"),
    education = recode(education, "1" = "less than high school", "2" = "high school equivalent", "3" = "more than high school")
  ) %>% 
  mutate(
    sex = factor(sex, levels = c("male", "female"), ordered = TRUE),
    education = factor(education, level = c("less than high school", "high school equivalent", "more than high school"), ordered = TRUE)
  )

acc_clean =
  read.csv("./data/nhanes_accel.csv") %>% 
  janitor::clean_names()
```

**Merge datasets**

```{r}
demo_acc = 
  left_join(demo_clean, acc_clean, by = "seqn")
```

**Table and plot for the number of men and women in education categories**

```{r}
sex_education =
  demo_acc %>% 
  group_by(sex, education) %>% 
  summarize(n = n()) %>% 
  pivot_wider(
    names_from = sex,
    values_from = n
  )

knitr::kable(sex_education)
```

```{r age_distribution_plot}
age_distribution_plot =
  demo_acc %>% 
  ggplot(aes(x = education, y = age, fill = sex)) +
  geom_boxplot(alpha = 0.5) +
  labs(
    x = "Education levels",
    y = "Age",
    title = "Age Distribution by Sex in Different Education Levels"
  ) +
  theme(legend.position = "right")

print(age_distribution_plot)
```
 > Comment
 
**Create total activity variable for each participants, and plot these total activities (y-axis) against age (x-axis).**
 
```{r acc_total_plot}
acc_total_plot =
  demo_acc %>% 
  mutate(total_activity = rowSums(select(., min1:min1440))) %>% 
  select("total_activity", everything()) %>% 
  ggplot(aes( x= age, y = total_activity, color = sex)) +
  geom_point(alpha = 0.5) +
  geom_smooth(se = FALSE) +
  labs(
    x = "Age",
    y = "Total activity",
    title = "Total activity Against Age by Sex and Education Level"
  ) +
  facet_grid(. ~ education)

print(acc_total_plot)
```

> Comment

**Three-panel plots for 24-hour activity time courses for each education level**

```{r allday_activity_plot}
allday_activity_plot =
  demo_acc %>% 
  pivot_longer(min1:min1440,
               names_to = "time_minute",
               values_to = "activity") %>%
  mutate(time_minute = as.numeric(gsub("min", "", time_minute))) %>% 
  ggplot(aes(x = time_minute, y = activity, color = sex)) +
  geom_point(alpha = 0.5) +
  geom_smooth(se = FALSE) +
  labs(
    x = "time(min)",
    y = "activity",
    color = "Sex",
    title = "24-hour Activity Time For Education Levels by Sex"
  ) +
  facet_grid(. ~ education)

allday_activity_plot
```

> Comment